\textbf{EECS 340: Assignment 1}

### 2020-01-27 assignment_1/article.tex
💄 Adjust format to have Problem 2(b) graph displayed continuously.
Co-Authored-By: bobbyCZT <zxc325@users.noreply.github.com>
♻️ Restructure Problem 2(d) with the loop invariant approach.

\documentclass[12pt]{article}
\usepackage{setspace}
\setstretch{1.1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\inlinecode}{\texttt}
\usepackage{etoolbox}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

@@

From \textit{Equation 2} and the known fact that $y = jd$, we may say that $d$ is also a common divisor of $x - y$ and $y$. By the defination of $gcd$, this means the upper bond of $d$ cannot be greater than $gcd(x-y, y)$. Thus we may conclude:
\begin{gather}
    gcd(x, y) = d \leq gcd(x-y, y) \nonumber \\
    \Rightarrow gcd(x, y) \leq gcd(x-y, y)
\end{gather}

@@

From \textit{Equation 4} and the known fact that $y = me$, we may say that $e$ is also a common divisor of $x$ and $y$. By the defination of $gcd$, this means the upper bond of $e$ cannot be greater than $gcd(x, y)$. Thus we may conclude:
\begin{gather}
    gcd(x-y, y) = e \leq gcd(x, y) \nonumber  \\
    \Rightarrow gcd(x-y, y) \leq gcd(x, y)
\end{gather}

@@

\newpage

@@

\subsection{(c)}
If $A[i] + B[j] \neq x$, then it is impossible to have any combination $(a, b)$ where \\ $a \in \{A[0], A[1], ..., A[i-1], A[i]\}$ and $b \in \{B[j], B[j+1], B[j+2], ..., B[n-1]\}$ to \\be $a + b = x$ for all \textbf{iterarted} $i, j$ within the \inlinecode{while} loop.

Assume we get $A[i] + B[j] = k$ within an interation of the \inlinecode{while} loop, we either have $k < x$ or $k > x$ (assume $A[i] + B[j] \neq x$ because otherwise the loop will be terminated and the problem will be solved). Due to the sorted nature of array $A, B$, we must have:
\begin{gather}
    \underbrace{A[0] \leq A[1] \leq ... \leq A[i-1]}_{a} \leq A[i]  \ \ \ \Rightarrow a \leq A[i]\\
    \underbrace{B[n-1] \geq ... \geq B[j+2] \geq B[j+1]}_{b} \geq B[j] \ \ \ \Rightarrow b \geq B[j]
\end{gather}

Thus, we may infer the followings from the above two \textit{Equations}:
\begin{gather}
    \text{if } k < x: \ a + B[j] \leq A[i] + B[j]  \ \ \ \Rightarrow a + B[j] < x\\
    \text{if } k > x: \ A[i] + b \geq A[i] + B[j]  \ \ \ \Rightarrow A[i] + b > x
\end{gather}

It is a bit unintuitive about the case of $a' + b'$ for $a' \in \{A[0], A[1], ..., A[i-1]\}$ and $b' \in \{B[j+1], B[j+2], ..., B[n-1]\}$. But as the value of index $i$ ($j$) travels in an incremental (decremental) fashion in the bond of $\mathbb{Z^+} \in [0, n)$; the index of $a'$ must be smaller than $i$, and the index of $b'$ must be greater than $j$. Thus, if there is any $a' + b' = x$, such loop would have been terminated immediately and we should not be able to reach to indexes $i, j$ in the \inlinecode{while} loop. Therefore, as we are now at the $i, j$ iteration of the \inlinecode{while} loop, this suggests there is no $(a', b')$ combination where $a' + b' = x$.

This together with \textit{Equation 10, 11}, we have proven the loop invariant as there will be no combination of $(a, b)$ where $a \in \{A[0], A[1], ..., A[i-1], A[i]\}$ and $b \in \{B[j], B[j+1], B[j+2], ..., B[n-1]\}$ to be $a + b = x$ for all \textbf{iterarted} $i, j$ within the \inlinecode{while} loop.

\subsection{(d)}
The worst case of this algorithm is when it returns \inlinecode{False}. In such case we wil have $i$ and $j$ travel the entire bond of $\mathbb{Z^+} \in [0, n)$. The loop will execute $2n$ times, thus the run time is $O(n)$.

>>>>

\documentclass[12pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\inlinecode}{\texttt}
\usepackage{etoolbox}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

@@

From \textit{Equation 2} and the known fact that $y = jd$, we may say that $d$ is also a common divisor of $x - y$ and $y$. By the defination of $gcd$, this means the upper bound of $d$ cannot be greater than $gcd(x-y, y)$. Thus we may conclude:
\begin{gather}
    gcd(x, y) = d \leq gcd(x-y, y) \nonumber \\
    \Rightarrow gcd(x, y) \leq gcd(x-y, y)
\end{gather}

@@

From \textit{Equation 4} and the known fact that $y = me$, we may say that $e$ is also a common divisor of $x$ and $y$. By the defination of $gcd$, this means the upper bound of $e$ cannot be greater than $gcd(x, y)$. Thus we may conclude:
\begin{gather}
    gcd(x-y, y) = e \leq gcd(x, y) \nonumber  \\
    \Rightarrow gcd(x-y, y) \leq gcd(x, y)
\end{gather}

@@

% \newpage

@@

\subsection{(c)}
\subsubsection{Loop invariant}
To enter the \inlinecode{while} loop with indexes $i, j$, it is impossible to have any combination of $(a, b)$ where $a \in \{A[0], A[1], ..., A[i-1]\}$ and $b \in \{B[j+1], B[j+2], ..., B[n-1]\}$ to be $a + b = x$.

\subsubsection{Initialization}
The loop invariant is held true during the initialization, since for $i = 0$ and $j = n-1$, we must have $a = \emptyset$ and $b = \emptyset$. Thus, it is impossible to have a combination of $(a, b)$ to be $a + b = x$.

Due to the sorted nature of array $A, B$, we must have:

\begin{gather}
    \underbrace{A[0] \leq A[1] \leq ... \leq A[i-1]}_{a} \leq A[i]  \ \ \ \Rightarrow a \leq A[i]\\
    \underbrace{B[n-1] \geq ... \geq B[j+2] \geq B[j+1]}_{b} \geq B[j] \ \ \ \Rightarrow b \geq B[j]
\end{gather}

Assume we get $A[i] + B[j] = k$ within an iteration of the \inlinecode{while} loop, we either have $k < x$ or $k > x$. In both cases, we may confidently say that there will be no $a + b = k$. As the index of $a$ must be smaller than $i$ and the index of $b$ must be greater than $j$. Thus, if there is any $a + b = x$, such loop would have been terminated immediately and we should not be able to reach to indexes $i, j$ in the \inlinecode{while} loop. Therefore, as we are now at the $i, j$ iteration of the \inlinecode{while} loop, this suggests there is no $(a, b)$ combination where $a + b = x$. By this, the loop invariant is held true for all cases within the \inlinecode{while} loop.\newline

Then, to proceed the loop, we may infer the following from \textit{Equation 8} and \textit{9}:

\begin{gather}
    \text{if } k < x: \ a + B[j] \leq A[i] + B[j] = k  \ \ \ \Rightarrow a + B[j] < x\\
    \text{if } k > x: \ A[i] + b \geq A[i] + B[j] = k  \ \ \ \Rightarrow A[i] + b > x
\end{gather}

As there is no $(a, b)$ combination available to form a $x$, we must increse the value of index $i$ in the case of $k < x$ to achieve a hopefully greather $k'$ in the next interation; and similarly, to decrese the value of index $j$ to achieve a hopefully lesser $k'$ in the next interation. Thus, the conditional statements in the pseudocode in \textbf{Section 2.1} are justified.

\subsection{(d)}
The worst case of this algorithm is when it returns \inlinecode{False}. In such case we will have $i$ and $j$ traveling the entire bound of $\mathbb{Z^+} \in [0, n)$. The loop will execute $2n$ times, thus the run time is $O(n)$.
